He-Man: A domain-specific language for lightweight concurrency

### 2011-12-13 report/report.tex
cleanup

>>>>

The He-Man runtime is written in C and provides a state machine based
thread scheduling model. In the view of the runtime, a thread consists
of some bookkeeping state, some private state, and the next
continuation function the thread is to run. The main part of the
runtime's state consists of the runqueue, which contains a list of all
currently runnable threads.

Continuation functions take a pointer to the thread information,
update the continuation field in the thread data, and return whether
the thread is still runnable; to block, a continuation function
registers itself as blocking on an event and returns false; to exit,
it frees the thread structure and returns false.
Running a thread, then, consists of repeatedly invoking the thread's
continuation function until it returns false. To prevent a
thread that does not need to block from starving the rest of the
system, after some number of invocations of the continuation
(currently 50), the thread is paused and put on the back of the
runqueue.

The He-Man runtime main loop is very simple, then: it loops
indefinitely, and on each iteration it removes the thread at the head
of the runqueue (if it exists) and runs it, and polls for IO events.

The runtime provides a blocking mechanism based on an event
abstraction. To the runtime, an event consists of a pointer to a
thread that is blocked on it and some arbitrary identifier (in
practice the file descriptor). To create an event for use with
nonblocking IO, an event is allocated, marked with the associated file
descriptor, and a new edge-triggered \t{epoll} event is created for the
file descriptor using \t{epoll\_ctl}, with the event structure as the
associated pointer data.
We use edge-triggered instead of level-triggered so that we can wait
on both read and write events for a given file descriptor without
having to adjust which we are waiting for at runtime, and because
He-Man's programming model makes it easy to meet the requirement that
blocking on a file descriptor is only done after an operation on the
descriptor has returned \t{EAGAIN}.
The actual polling for IO events is done by calling \t{epoll}. For each
event that has been returned by epoll, if a thread is waiting on the
event, the thread is added to the runqueue. If the the runqueue is
empty, the call to epoll is blocking, otherwise it is called with a
timeout of zero seconds, to simply poll for events.

The runtime also has support for Linux Kernel Asynchronous IO for file
IO, using a Linux \t{eventfd} to allow \t{epoll} to wait on AIO
events. This support is currently unused, since it was not actually
necessary to acheive acceptable performance for the web server.
